### 2024-10-04 paper/header.tex

% **************GENERATED FILE, DO NOT EDIT**************

\title{DisjunctiveProgramming.jl: Generalized Disjunctive Programming Models and Algorithms for JuMP}

\author[1]{Hector D. Perez}
\author[1]{Shivank Joshi}
\author[1]{Ignacio E. Grossmann}
\affil[1]{Carnegie Mellon University}

\keywords{JuMP, Mathematical Optimization, Generalized Disjunctive Programming}

\hypersetup{
pdftitle = {DisjunctiveProgramming.jl: Generalized Disjunctive Programming Models and Algorithms for JuMP},
pdfsubject = {JuliaCon 2019 Proceedings},
pdfauthor = {Hector D. Perez, Shivank Joshi, Ignacio E. Grossmann},
pdfkeywords = {JuMP, Mathematical Optimization, Generalized Disjunctive Programming},
}